Advancements in Moltres for Time-Dependent Multiphysics Molten Salt Reactor Modeling

### 2024-11-19 msre.tex
add ray effect subsection

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{msre-quarter-no-rod-power}
  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model.
  This figure depicts the 1/8th-core distribution due to symmetry about the diagonal.}
  \label{fig:1/4-no-rod}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\columnwidth]{msre-quarter-rod-power}
  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model.
  This figure depicts the 1/8th-core distribution due to symmetry about the diagonal.}
  \label{fig:1/4-rod}
\end{figure}

@@

>>>>

\begin{figure}[p]
  \centering
  \includegraphics[width=\columnwidth]{msre-quarter-no-rod-power}
  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model.
  This figure depicts the 1/8th-core distribution due to symmetry about the diagonal.}
  \label{fig:1/4-no-rod}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=\columnwidth]{msre-quarter-rod-power}
  \caption{Normalized channel fission rate distribution of the 2-D \gls{MSRE} quarter-core model.
  This figure depicts the 1/8th-core distribution due to symmetry about the diagonal.}
  \label{fig:1/4-rod}
\end{figure}

@@

\subsubsection{$S_N$ Discretization \& Ray Effects}

Angular resolution in 2-D (and later 3-D) flux simulations is important for accurately capturing
the angular dependence in the neutron angular flux. This subsection compares the $k_\text{eff}$ and
neutron scalar flux distribution solutions from the hybrid method with the $S_6$, $S_8$, $S_{10}$,
and $S_{12}$ methods for the 2-D full-core with no rods inserted. Table
\ref{table:2d-sn-convergence} shows the $k_\text{eff}$ estimates from those simulation results.
The $k_\text{eff}$ varies by less than 1 pcm, thereby indicating that the angular discretization
order used by the hybrid method has little impact on the integral reactor parameters.

Figures \ref{fig:ray-effect-12} and \ref{fig:ray-effect-34} show the first four neutron group flux
distributions from the hybrid method in the central reactor region consisting of three
air-filled thimbles and a homogenized sample basket (bottom right). The geometry corresponds
directly to Figure \ref{fig:full-geom-closeup} without the three control rods. Group 1 fluxes
in the thimble form a distinctive ray effect artifacts arising from the regular geometrical shapes and
insufficient angular resolution. The artifacts become less distinct with increasing angular
discretization levels. They also diminish with increasing neutron energy groups due to shorter
neutron mean free paths leading to more diffusive behavior.

By Group 3 and 4, the artifacts are indiscernible to the eye across all angular discretization
orders. Given that $k_\text{eff}$ is heavily weighted towards the later (slower) energy groups due
to larger fission cross sections, it is clear why Table \ref{table:2d-sn-convergence} shows little
differences in $k_\text{eff}$ across the angular discretization orders. Therefore, $S_8$ is
adequate for control rod worth calculations in the \gls{MSRE}. 
Further investigations into the angular discretization level may be warranted when applying the
hybrid method to reactors with more air-filled regions such as high-temperature gas-cooled
reactors.

\begin{table}[t]
  \centering
  \caption{Hybrid method $k_\text{eff}$ estimates for the 2-D full-core model with zero rods
  inserted and $S_N$ schemes from $S_6$ to $S_{12}$.}
  \begin{tabular}{l S}
    \toprule
    Angular discretization & {$k_\text{eff}$} \\
    \midrule
    $S_6$ & 1.108414 \\
    $S_8$ & 1.108423 \\
    $S_{10}$ & 1.108422 \\
    $S_{12}$ & 1.108420 \\
    \bottomrule
  \end{tabular}
  \label{table:2d-sn-convergence}
\end{table}

\begin{figure}[p]
  \small
  \centering
  \includegraphics[width=0.7\columnwidth]{ray-effect-12}
  \caption{Group 1 and 2 neutron flux distributions in the hybrid $S_N$-diffusion method correction
  region with $S_6$, $S_8$, $S_{10}$, \& $S_{12}$. Ray effects diminish with increasing
  angular fidelity, but still persist with $S_{12}$ in group 1 and 2.}
  \label{fig:ray-effect-12}
\end{figure}

\begin{figure}[p]
  \small
  \centering
  \includegraphics[width=0.7\columnwidth]{ray-effect-34}
  \caption{Group 3 and 4 neutron flux distributions in the hybrid $S_N$-diffusion method correction
  region with $S_6$, $S_8$, $S_{10}$, \& $S_{12}$. Ray effects are barely visible in group
  3 and 4 due to shorter neutron mean free paths.}
  \label{fig:ray-effect-34}
\end{figure}
